---
title: "explore_gastruloid_data"
author: "Sophie Seidel"
date: "2023-04-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "~/Projects/typewriter_analysis/")
## load up the packages we will need:  (uncomment as required)
library(ggplot2)
library(reshape2)
## ---------------------------

## load up our functions into memory
source("src/useful_scripts_across_categories.R")
```

```{r io}
# output dir
plot_dir = "results/preliminary_gastruloid/0_explore_dataset/"

# input file
filtered_dat_file = "data/preliminary_gastruloid/mGASv2_Lane2_CellByTape_filtered_for_8barcodes.RDS"
filtered_dat = readRDS(filtered_dat_file)
```

```{r different_sequences}
# Reduce to cells with different sequences
filtered_dat_unique_cells = unique(filtered_dat[, 2:7])
```


```{r n edits all seqs}
filtered_dat$nrOfEdits = get_nr_of_edits(filtered_dat[, 1:7], unedited_symbol = "None")

g = ggplot(filtered_dat, aes(x=nrOfEdits)) +
  facet_wrap(~ TargetBC)+
  geom_histogram()+
  theme_minimal() +
  ggtitle("Number of edits per targetBC", subtitle = "For all sequences")
g
```

Next, we only keep sequences that are unique.


```{r n edits unique seqs}
## plot number of edits at different targetBCs
filtered_dat_unique_cells$nrOfEdits = get_nr_of_edits(filtered_dat_unique_cells, unedited_symbol = "None")

g = ggplot(filtered_dat_unique_cells, aes(x=nrOfEdits)) +
  facet_wrap(~ TargetBC)+
  geom_histogram()+
  theme_minimal()+
  ggtitle("Number of edits per targetBC", subtitle = "For unique sequences")
g
```


We can make two observations from the above plots:

 - the distribution of the number of edits shifts when filtering cell with identical sequences. These difference are most likely a consequence of the population process.
 - Some targets (e.g. TAACGAAGATTT, TTGATAACGTGA) show notable difference to the edit distributions in other targets. Hence, we will use a clock per target.
 
