---
title: "Gastruloid exploratory analysis"
output: html_document
---
This is meant as a document to explore features of the gastruloid Typewriter data, preliminarily.
Reading in data and removing last site:
```{r setup, include=FALSE}

edit_table_by_6 = read.csv("/Users/azwaans/typewriter_analysis/data/mGASv2_Lane2_CellByTape_10X_bamExtractV2_t3_collapse.csv", stringsAsFactors = F, header = T,na.strings=c("","NA"))

edit_table_by_5 = edit_table_by_6[,1:7]

```


Basic properties of the dataset: 

Total number of cell sampled:
```{r ,include=FALSE}
length(unique(edit_table_by_5$Cell))
```


Number of unique TargetBC recovered:
```{r ,include=FALSE}
length(unique(edit_table_by_5$TargetBC))
```





Filtering out the 8 TargetBCs specified by Choi:
```{r ,include=FALSE}
filtered_dataset <- c()
for(i in filters$TargetBC) {
  filtered_dataset <- rbind(filtered_dataset,edit_table_by_5[which(edit_table_by_5$TargetBC == i),])
}

filtered_by_targetBC <- filtered_dataset

frequent_target_bcs <- filters$TargetBC
##further filter by cell that have ALL 8. 
full_filter<- filtered_dataset
for (cell in unique(full_filter$Cell)){

  print(cell)

  cell_target_bcs = full_filter[which(full_filter$Cell == cell), "TargetBC"]
  keep_cell = all(is.element(el = frequent_target_bcs, set = cell_target_bcs))

  if (! keep_cell){
    full_filter = full_filter[which(!(full_filter$Cell == cell)), ]
  }
}


```


Total number of cell sampled:
```{r ,include=FALSE}
length(unique(edit_table_by_5$Cell))
```
Number of unique TargetBC recovered in the full dataset:
```{r }
length(unique(edit_table_by_5$TargetBC))
```

Which are the 8 TargetBCs to filter specified by Choi:
```{r }
filters = read.csv("/Users/azwaans/typewriter_analysis/data/mGASv2_TargetBC_selected8.csv", header = T,na.strings=c("","NA"))
filters$TargetBC
```

Total number of cells that have some of the 8 chosen targetBCs 
```{r }
length(unique(filtered_by_targetBC$Cell))
```
Total number of cells that have all of the 8 chosen targetBCs 
```{r }
length(unique(full_filter$Cell))
```
How many unique sequences/targetBC? 
```{r }
##first, needs to be converted into sequences such that they can be compared. 
# 1st remove all trinucleotides and convert NAs with NA
sequences_filtered <- full_filter
for(i in 3:7) {
  sequences_filtered[,i] <- substring(sequences_filtered[,i], 1,3)
  sequences_filtered[which(is.na(sequences_filtered[,i])),i] <- "NA"

}

# 2nd, map these to an integer:

# get this from the frequencies:
bulk_insert_count <- data.frame(table(unlist(sequences_filtered[,3:7]))) %>% arrange(desc(Freq))
bulk_insert_count$Var1 <- substring(bulk_insert_count$Var1, 1,3)
bulk_insert_count$codes <- as.character(seq(0,42))
bulk_insert_count <- data.frame(edit=bulk_insert_count$Var1, code=as.character(seq(0,42)))
#bulk_insert_count <- rbind(c("NA",0),bulk_insert_count)
pick_code <- function(edit,code_map) {
  code <- code_map[which(code_map$edit==edit),2]
  return(code)

}


# convert these trinucleotides to integer format:
for(i in 3:7) {
  sequences_filtered[,i] <- unlist(lapply(sequences_filtered[,i], function(x) {pick_code(x,bulk_insert_count)}))

}

# concatenate them, adding commas:
sequences_filtered$beast_seq <- apply(sequences_filtered[,3:7],1,function(x) {str_flatten(x,collapse = ",")})

##simple unique ones
unique <- c()
for(i in unique(sequences_filtered$TargetBC)) {

  unique <-c(unique,length(unique(sequences_filtered[sequences_filtered$TargetBC == i,"beast_seq"])))
  
}

##count unedited bcodes
unedited <- c()
for(i in unique(sequences_filtered$TargetBC)) {
  unedited <- c(unedited,length(which(sequences_filtered[sequences_filtered$TargetBC == i,"beast_seq"] == "0,0,0,0,0" )))
  
}
#those are all fully edited sequences
all_fully_edited <- sequences_filtered[which(!unlist(lapply(sequences_filtered$beast_seq,function(x) {stringr::str_ends(x,pattern ='0')}))),]
unique_fully_edited <- c()
for(i in unique(all_fully_edited$TargetBC)) {
  unique_fully_edited <- c(unique_fully_edited,length(unique(all_fully_edited[all_fully_edited$TargetBC == i,"beast_seq"])))
  
}

unique_by_type <- data.frame(cbind(unedited,unique,unique_fully_edited))
unique_by_type <- data.frame(cbind(targetBC=unique(all_fully_edited$TargetBC),unique_by_type))


unique_by_type <- pivot_longer(unique_by_type,!targetBC,names_to = "Type",values_to = "count")

ggplot(unique_by_type, aes(x=targetBC,y=count,fill=Type,)) + geom_bar(stat = "identity",position=position_dodge()) + theme(axis.text.x =  element_text(angle = 90, vjust = 0.5, hjust=1))
```



Plotting basic stats:
Trinucleotide sequences:
```{r}
filtered_by_targetBC[filtered_by_targetBC=='None'] <- NA
bulk_insert_count <- data.frame(table(unlist(filtered_by_targetBC[,3:7]))) %>% arrange(desc(Freq))
bulk_insert_count$Var1 <- substring(bulk_insert_count$Var1, 1,3)
bulk_insert_count$Var1[1] <- "NA"
bulk_insert_count$Var1 <- factor(bulk_insert_count$Var1,levels=bulk_insert_count$Var1)
plot_filtered <- ggplot(data=bulk_insert_count, aes(x=Var1,y=Freq,fill=Var1)) + geom_col() + xlab("Insert sequence") + ylab("Total counts") +  theme(legend.title = element_blank(),axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1))

full_filter[full_filter=='None'] <- NA
bulk_insert_count <- data.frame(table(unlist(full_filter[,3:7]))) %>% arrange(desc(Freq))
bulk_insert_count$Var1 <- substring(bulk_insert_count$Var1, 1,3)
bulk_insert_count$Var1[1] <- "NA"
bulk_insert_count$Var1 <- factor(bulk_insert_count$Var1,levels=bulk_insert_count$Var1)
plot_filtered_full <- ggplot(data=bulk_insert_count, aes(x=Var1,y=Freq,fill=Var1)) + geom_col() + xlab("Insert sequence") + ylab("Total counts") +  theme(legend.title = element_blank(),axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1))

cowplot::plot_grid(plot_filtered + theme(legend.position = "None"),plot_filtered_full + theme(legend.position = "None"))

```


```{r}
bulk_insert_count <- data.frame(table(unlist(filtered_by_targetBC[,3:7]))) %>% arrange(desc(Freq))
bulk_insert_count$Var1 <- substring(bulk_insert_count$Var1, 1,3)
bulk_insert_count$Var1 <- factor(bulk_insert_count$Var1,levels=bulk_insert_count$Var1)
plot_filtered <- ggplot(data=bulk_insert_count, aes(x=Var1,y=Freq/sum(Freq),fill=Var1)) + geom_col() + xlab("Insert sequence") + ylab("Frequency")  +  theme(legend.title = element_blank(),axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1))

bulk_insert_count <- data.frame(table(unlist(full_filter[,3:7]))) %>% arrange(desc(Freq))
bulk_insert_count$Var1 <- substring(bulk_insert_count$Var1, 1,3)
bulk_insert_count$Var1 <- factor(bulk_insert_count$Var1,levels=bulk_insert_count$Var1)
plot_filtered_full <- ggplot(data=bulk_insert_count, aes(x=Var1,y=Freq/sum(Freq),fill=Var1)) + geom_col() + xlab("Insert sequence") + ylab("Frequency")  +  theme(legend.title = element_blank(),axis.text.x = element_text(angle = 45, vjust = 0.5, hjust=1))

cowplot::plot_grid(plot_filtered + theme(legend.position = "None"),plot_filtered_full + theme(legend.position = "None"))

```

Looking at insert distributions per site: there seems to be some bias for AAA inserted at site 2?
```{r}
insert_count_site <- data.frame(filtered_by_targetBC[,3:7]) %>% gather(site, sequence,1:5)
insert_count_site$sequence <- substring(insert_count_site$sequence, 1,3)
plot_filtered <- ggplot(insert_count_site, aes(x=site,fill=sequence)) + geom_histogram(stat="count") + xlab("Site") + ylab("Total counts") + theme(legend.title = element_blank())

insert_count_site <- data.frame(full_filter[,3:7]) %>% gather(site, sequence,1:5)
insert_count_site$sequence <- substring(insert_count_site$sequence, 1,3)
plot_filtered_full <- ggplot(insert_count_site, aes(x=site,fill=sequence)) + geom_histogram(stat="count") + xlab("Site") + ylab("Total counts") + theme(legend.title = element_blank())

cowplot::plot_grid(plot_filtered + theme(legend.position = "None") ,plot_filtered_full + theme(legend.position = "None"))

```
```{r}
insert_count_site <- data.frame(filtered_by_targetBC[,3:7]) %>% gather(site, sequence,1:5)
insert_count_site$sequence <- substring(insert_count_site$sequence, 1,3)
insert_count_site <- insert_count_site[- which(is.na(insert_count_site$sequence)),]

#testing whether insert distributions are independent of sites?
plot_filtered <- ggplot(insert_count_site, aes(x=site,fill=sequence)) + geom_bar() + xlab("Site") + ylab("Total counts")+ theme(legend.title = element_blank())

insert_count_site <- data.frame(full_filter[,3:7]) %>% gather(site, sequence,1:5)
insert_count_site$sequence <- substring(insert_count_site$sequence, 1,3)
insert_count_site <- insert_count_site[- which(is.na(insert_count_site$sequence)),]

#testing whether insert distributions are independent of sites?
plot_filtered_full <- ggplot(insert_count_site, aes(x=site,fill=sequence)) + geom_bar() + xlab("Site") + ylab("Total counts")+ theme(legend.title = element_blank())

cowplot::plot_grid(plot_filtered + theme(legend.position = "None") ,plot_filtered_full + theme(legend.position = "None"))

```

```{r}
insert_count_site <- data.frame(filtered_by_targetBC[,3:7]) %>% gather(site, sequence,1:5)
insert_count_site$sequence <- substring(insert_count_site$sequence, 1,3)
insert_count_site <- insert_count_site[- which(is.na(insert_count_site$sequence)),]

#testing whether insert distributions are independent of sites?
plot_filtered <- ggplot(insert_count_site, aes(x=site)) + geom_bar() + xlab("Site") + ylab("Total counts")+ theme(legend.title = element_blank())

insert_count_site <- data.frame(full_filter[,3:7]) %>% gather(site, sequence,1:5)
insert_count_site$sequence <- substring(insert_count_site$sequence, 1,3)
insert_count_site <- insert_count_site[- which(is.na(insert_count_site$sequence)),]

#testing whether insert distributions are independent of sites?
plot_filtered_full <- ggplot(insert_count_site, aes(x=site)) + geom_bar() + xlab("Site") + ylab("Total counts")+ theme(legend.title = element_blank())

cowplot::plot_grid(plot_filtered + theme(legend.position = "None") ,plot_filtered_full + theme(legend.position = "None"))

```
```{r}
insert_freq_site  <- data.frame(filtered_by_targetBC[,3:7]) %>% gather(site, sequence,1:5)
insert_freq_site$sequence <- substring(insert_freq_site$sequence, 1,3)
insert_freq_site <- count(insert_freq_site, insert_freq_site$site, insert_freq_site$sequence)
colnames(insert_freq_site) <- c("site","sequence","count")
plot_filtered <- ggplot(insert_freq_site, aes(fill=sequence, y=count, x=site)) +
  geom_bar(position="fill", stat="identity") + xlab("Site") + ylab("Frequency") + theme(legend.title = element_blank())

insert_freq_site  <- data.frame(full_filter[,3:7]) %>% gather(site, sequence,1:5)
insert_freq_site$sequence <- substring(insert_freq_site$sequence, 1,3)
insert_freq_site <- count(insert_freq_site, insert_freq_site$site, insert_freq_site$sequence)
colnames(insert_freq_site) <- c("site","sequence","count")
plot_filtered_full <- ggplot(insert_freq_site, aes(fill=sequence, y=count, x=site)) +
  geom_bar(position="fill", stat="identity") + xlab("Site") + ylab("Frequency") + theme(legend.title = element_blank())

cowplot::plot_grid(plot_filtered + theme(legend.position = "None") ,plot_filtered_full + theme(legend.position = "None"))
```


```{r}
insert_freq_site  <- data.frame(filtered_by_targetBC[,3:7]) %>% gather(site, sequence,1:5)
insert_freq_site$sequence <- substring(insert_freq_site$sequence, 1,3)
insert_freq_site <- insert_freq_site[- which(is.na(insert_freq_site$sequence)),]
insert_freq_site <- count(insert_freq_site, insert_freq_site$site, insert_freq_site$sequence)
colnames(insert_freq_site) <- c("site","sequence","count")
plot_filtered <- ggplot(insert_freq_site, aes(fill=sequence, y=count, x=site)) +
  geom_bar(position="fill", stat="identity") + xlab("Site") + ylab("Frequency") + theme(legend.title = element_blank())

insert_freq_site  <- data.frame(full_filter[,3:7]) %>% gather(site, sequence,1:5)
insert_freq_site$sequence <- substring(insert_freq_site$sequence, 1,3)
insert_freq_site <- insert_freq_site[- which(is.na(insert_freq_site$sequence)),]
insert_freq_site <- count(insert_freq_site, insert_freq_site$site, insert_freq_site$sequence)
colnames(insert_freq_site) <- c("site","sequence","count")
plot_filtered_full <-ggplot(insert_freq_site, aes(fill=sequence, y=count, x=site)) +
  geom_bar(position="fill", stat="identity") + xlab("Site") + ylab("Frequency") + theme(legend.title = element_blank())

cowplot::plot_grid(plot_filtered + theme(legend.position = "None") ,plot_filtered_full + theme(legend.position = "None"))
```
```{r}
insert_count_targBC <- data.frame(filtered_by_targetBC[,c(2,3,4,5,6,7)]) %>% gather(targetBC, sequence,2:6)
insert_count_targBC <- insert_count_targBC[, c(1,3)]
insert_count_targBC$sequence <- substring(insert_count_targBC$sequence, 1,3)
plot_filtered <- ggplot(insert_count_targBC, aes(x=TargetBC,fill=sequence)) + geom_bar() + theme(axis.text.x = element_text(angle = 90)) +
  xlab("TargetBC") + ylab("Count") + theme(legend.title = element_blank())

insert_count_targBC <- data.frame(full_filter[,c(2,3,4,5,6,7)]) %>% gather(targetBC, sequence,2:6)
insert_count_targBC <- insert_count_targBC[, c(1,3)]
insert_count_targBC$sequence <- substring(insert_count_targBC$sequence, 1,3)
plot_filtered_full <- ggplot(insert_count_targBC, aes(x=TargetBC,fill=sequence)) + geom_bar() + theme(axis.text.x = element_text(angle = 90)) +
  xlab("TargetBC") + ylab("Count") + theme(legend.title = element_blank())

cowplot::plot_grid(plot_filtered + theme(legend.position = "None") ,plot_filtered_full + theme(legend.position = "None"))
```
```{r}
insert_count_targBC <- data.frame(filtered_by_targetBC[,c(1,2,3,4,5,6,7)]) %>% gather(targetBC, sequence,3:7)
insert_count_targBC <- insert_count_targBC[- which(is.na(insert_count_targBC$sequence)),]
insert_count_targBC$sequence <- substring(insert_count_targBC$sequence, 1,3)
ggplot(insert_count_targBC, aes(x=TargetBC,fill=sequence)) + geom_bar() + theme(axis.text.x = element_text(angle = 90)) +
  xlab("TargetBC") + ylab("Count") + theme(legend.title = element_blank())

insert_count_targBC <- data.frame(full_filter[,c(1,2,3,4,5,6,7)]) %>% gather(targetBC, sequence,3:7)
insert_count_targBC <- insert_count_targBC[- which(is.na(insert_count_targBC$sequence)),]
insert_count_targBC$sequence <- substring(insert_count_targBC$sequence, 1,3)
ggplot(insert_count_targBC, aes(x=TargetBC,fill=sequence)) + geom_bar() + theme(axis.text.x = element_text(angle = 90)) +
  xlab("TargetBC") + ylab("Count") + theme(legend.title = element_blank())

cowplot::plot_grid(plot_filtered + theme(legend.position = "None") ,plot_filtered_full + theme(legend.position = "None"))
```
```{r}
insert_count_targBC <- data.frame(filtered_by_targetBC[,c(1,2,3,4,5,6,7)]) %>% gather(targetBC, sequence,3:7)
insert_count_targBC <- insert_count_targBC[- which(is.na(insert_count_targBC$sequence)),]
insert_count_targBC$sequence <- substring(insert_count_targBC$sequence, 1,3)
cell_count_targBC <- count(insert_count_targBC, insert_count_targBC$TargetBC,insert_count_targBC$Cell)
unique_cells <- c()
for( i in unique(cell_count_targBC$`insert_count_targBC$TargetBC`)) {
  unique_cells <- c(unique_cells,length(unique(cell_count_targBC[cell_count_targBC$`insert_count_targBC$TargetBC` == i,2])))

}

cells_per_targetBC <- data.frame(targetBC= unique(cell_count_targBC$`insert_count_targBC$TargetBC`),unique_cells = unique_cells )
plot_filtered <- ggplot(cells_per_targetBC,aes(x=targetBC,y=unique_cells)) + geom_bar(stat = "identity") + theme(axis.text.x = element_text(angle = 90)) +
  xlab("TargetBC") + ylab("Number of cells")

insert_count_targBC <- data.frame(full_filter[,c(1,2,3,4,5,6,7)]) %>% gather(targetBC, sequence,3:7)
insert_count_targBC <- insert_count_targBC[- which(is.na(insert_count_targBC$sequence)),]
insert_count_targBC$sequence <- substring(insert_count_targBC$sequence, 1,3)
cell_count_targBC <- count(insert_count_targBC, insert_count_targBC$TargetBC,insert_count_targBC$Cell)
unique_cells <- c()
for( i in unique(cell_count_targBC$`insert_count_targBC$TargetBC`)) {
  unique_cells <- c(unique_cells,length(unique(cell_count_targBC[cell_count_targBC$`insert_count_targBC$TargetBC` == i,2])))

}

cells_per_targetBC <- data.frame(targetBC= unique(cell_count_targBC$`insert_count_targBC$TargetBC`),unique_cells = unique_cells )
plot_filtered_full <- ggplot(cells_per_targetBC,aes(x=targetBC,y=unique_cells)) + geom_bar(stat = "identity") + theme(axis.text.x = element_text(angle = 90)) +
  xlab("TargetBC") + ylab("Number of cells")

cowplot::plot_grid(plot_filtered + theme(legend.position = "None") ,plot_filtered_full + theme(legend.position = "None"))
```
TAACGAAGATTT seems to be the bottleneck for the number of cells containing all TargetBCs


```{r}
insert_freq_targBC <- data.frame(filtered_by_targetBC[,c(1,2,3,4,5,6,7)]) %>% gather(site, sequence,3:7)
insert_freq_targBC$sequence <- substring(insert_freq_targBC$sequence, 1,3)
insert_freq_targBC <- count(insert_freq_targBC, insert_freq_targBC$TargetBC, insert_freq_targBC$sequence)
colnames(insert_freq_targBC) <- c("target_BC","sequence","count")
plot_filtered <- ggplot(insert_freq_targBC, aes(fill=sequence, y=count, x=target_BC)) + theme(axis.text.x = element_text(angle = 90)) +
  geom_bar(position="fill", stat="identity") +  theme(legend.title = element_blank()) +
  xlab("TargetBC") + ylab("Frequency")  + theme(legend.title = element_blank())

insert_freq_targBC <- data.frame(full_filter[,c(1,2,3,4,5,6,7)]) %>% gather(site, sequence,3:7)
insert_freq_targBC$sequence <- substring(insert_freq_targBC$sequence, 1,3)
insert_freq_targBC <- count(insert_freq_targBC, insert_freq_targBC$TargetBC, insert_freq_targBC$sequence)
colnames(insert_freq_targBC) <- c("target_BC","sequence","count")
plot_filtered_full <- ggplot(insert_freq_targBC, aes(fill=sequence, y=count, x=target_BC)) + theme(axis.text.x = element_text(angle = 90)) +
  geom_bar(position="fill", stat="identity") +  theme(legend.title = element_blank()) +
  xlab("TargetBC") + ylab("Frequency")  + theme(legend.title = element_blank())

cowplot::plot_grid(plot_filtered + theme(legend.position = "None") ,plot_filtered_full + theme(legend.position = "None"))
```
Compared to the cell culture data, we observe much more heterogeneity in edit distributions per target BCs. 
Could it be that this is due to much less editing on average? 
ie. the more edited a Target BC, the more the distribution should tend to the true? 
The one ending in TTTG is mmost edited. 
Is this better in the subsample? 





```{r}
insert_freq_targBC <- data.frame(filtered_by_targetBC[,c(2,3,4,5,6,7)]) %>% gather(targetBC, sequence,2:6)
insert_freq_targBC <- insert_freq_targBC[ which(! is.na(insert_freq_targBC$sequence)),]
insert_freq_targBC$sequence <- substring(insert_freq_targBC$sequence, 1,3)
insert_freq_targBC <- count(insert_freq_targBC, insert_freq_targBC$TargetBC, insert_freq_targBC$sequence)
colnames(insert_freq_targBC) <- c("target_BC","sequence","count")
plot_filtered <- ggplot(insert_freq_targBC, aes(fill=sequence, y=count, x=target_BC)) +
  geom_bar(position="fill", stat="identity") + theme(axis.text.x = element_text(angle = 90)) +
  xlab("TargetBC") + ylab("Frequency") + theme(legend.title = element_blank())

insert_freq_targBC <- data.frame(full_filter[,c(2,3,4,5,6,7)]) %>% gather(targetBC, sequence,2:6)
insert_freq_targBC <- insert_freq_targBC[ which(! is.na(insert_freq_targBC$sequence)),]
insert_freq_targBC$sequence <- substring(insert_freq_targBC$sequence, 1,3)
insert_freq_targBC <- count(insert_freq_targBC, insert_freq_targBC$TargetBC, insert_freq_targBC$sequence)
colnames(insert_freq_targBC) <- c("target_BC","sequence","count")
plot_filtered_full <-ggplot(insert_freq_targBC, aes(fill=sequence, y=count, x=target_BC)) +
  geom_bar(position="fill", stat="identity") + theme(axis.text.x = element_text(angle = 90)) +
  xlab("TargetBC") + ylab("Frequency") + theme(legend.title = element_blank())

cowplot::plot_grid(plot_filtered + theme(legend.position = "None") ,plot_filtered_full + theme(legend.position = "None"))
```


Getting the heatmap of conditional probabilities for the fully filtered dataset
```{r}
full_filter[is.na(full_filter)] <- "NA"
all_inserts <- unique(unlist(full_filter[,3:7]))
contiguous_inserts <- matrix(0,nrow = length(all_inserts),ncol = length(all_inserts))
rownames(contiguous_inserts) = all_inserts
colnames(contiguous_inserts) = all_inserts

#counting manually the joint probs.
for(i in 1:(length(full_filter$Cell))) {
  print(paste0("i ", i))
  for (j in 3:6) {
      print(paste0("cell ", j))
    for (a in 1:length(all_inserts)) {
      
      for (b in 1:length(all_inserts)) {
        if((full_filter[i,j] == all_inserts[a]) && (full_filter[i,j+1] == all_inserts[b])) {
          contiguous_inserts[a,b] = contiguous_inserts[a,b] + 1
        }
      }
    }
  }



}
write.csv(contiguous_inserts,"results/preliminary_gastruloid/0_explore_dataset/contiguous_gastruloids.csv")
#converting to conditional probs
bigram_matrix <- data.frame(contiguous_inserts)
bigram_matrix <- bigram_matrix[-which(rownames(bigram_matrix) == "None"), -which(colnames(bigram_matrix) == "None")]
#divide rows by their rowsums:
bigram_matrix <- apply(bigram_matrix, 1, function(x) {x=x/sum(x)})
dim(bigram_matrix)
#very dirty way to melt the matrix to a 3 column thing
count <- c()
position1 <- c()
for( i in rownames(bigram_matrix)) {

  position1 <- c(position1,rep(i,42))

}
position1 <- unlist(position1)
position2 <- rep(colnames(bigram_matrix),42)
counts <- as.vector(bigram_matrix)

#clean up insert symbols
position1 <- substring(position1, 1,3)
position2 <- substring(position2, 1,3)

heatmap_data <- data.frame(pos1 = position1, pos2=position2, count = counts)
View(heatmap_data)
ggplot(heatmap_data, aes(x=pos1, y=pos2, fill=count)) +
  geom_tile() + xlab("Position 1") + ylab("Position 2") + geom_tile(color = "black") +
  scale_fill_gradient2(low = "#075AFF",
                       mid = "#FFFFCC",
                       high = "#FF0000") + labs(fill="P(Position2|Position1)") + coord_fixed()

```
 